### 2019-08-18 chapters/4-approach.tex
Add examples of the obtained results
Finish experiments section and conclusion

\subsection{Parsing}
Before comparison, declaration files need to be parsed so that both of them share the same structure. It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \citep{ts-compiler-api}.

@@

\section{Type Inference}
Type inference for JavaScript is a topic that was deeply investigated. Nevertheless, it was not included in the implementation. Types were directly imported from the types at run-time. For example, if a function \mintinline{text}{f(input)} was invoked as \mintinline{text}{f("hello")}, argument \mintinline{text}{input} would be declared as \mintinline{text}{string} in the declaration file.
The refinement of the overall architecture of the generation of TypeScript declaration files was prioritized over type inference accuracy. Extracting the code examples from the repositories proved to be a solid and pragmatic way of capturing the developer's intention. Using the types imported from those examples allowed to focus on creating an end-to-end solution while still obtaining declaration files with reasonable types. It is worth noting that this would not have been the case if randomly generated test cases were used.
The discussed approach on how to deal with type inference in JavaScript is presented in detail below.

>>>>

\subsection{Parsing} \label{sec:approach-parsing}
Before comparison, declaration files need to be parsed so that both of them share the same structure. It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \citep{ts-compiler-api}.

@@

\section{Type Inference} \label{sec:approach-type-inference}
Type inference for JavaScript is a topic that was deeply investigated. Nevertheless, it was not included in the implementation. Types were directly imported from the types at run-time. For example, if a function \mintinline{text}{f(input)} was invoked as \mintinline{text}{f("hello")}, argument \mintinline{text}{input} would be declared as \mintinline{text}{string} in the declaration file.
The refinement of the overall architecture of the generation of TypeScript declaration files was prioritized over type inference accuracy. Extracting the code examples from the repositories proved to be a solid and pragmatic way of capturing the developer's intention. Using the types imported from those examples allowed to focus on creating an end-to-end solution while still obtaining declaration files with reasonable types. It is worth noting that this would not have been the case if randomly generated test cases were used.
The discussed approach on how to deal with type inference in JavaScript is presented in detail below.

### 2019-08-18 chapters/5-experiments.tex
Finish experiments section and conclusion

\chapter{Experiments}\label{chap:experiments}

Experiments often include benchmarks.
You need to describe you benchmarking setup: parameters of the machine used (architecture, memory, etc), parameters of the software used (OS version, libraries used and their version), the benchmark programs.
The programs need not be part of the thesis, but they should be described and available separately.

If you measure run times and evaluate your findings statistically, then there is great potential for errors.
Consult SIGPLAN's checklist for empirical evaluations \url{http://www.sigplan.org/Resources/EmpiricalEvaluation/} for advice.

\input{figures/experiments/figExample}
\input{figures/experiments/tabAccuracy}

\section{Definitely Typed Modules}

\section{Code Examples Extraction}
\subsection{Repositories URL}
\subsection{Readme Files}
\subsection{Code Examples}

\section{Code Execution}

\section{Run-time Information Gathering}

\section{Declaration Files Generation}

\section{Evaluation}

\section{JS Operators Usage}
\todo{Agregar que es una matriz simetrica. Que era para hacer un analisis mas global. Para algunos operators se puede inferir que el orden de los tipos sí afecta. Por ejemplo para instance of donde obviamente importa el orden. Pero para == por ejemplo generalmente se pone a == null y no null == a}.
\input{figures/experiments/operators/relational-operators/relationalOperators.tex}
\input{figures/experiments/operators/additive-operators/additiveOperators.tex}
\input{figures/experiments/operators/multiplicative-operators/multiplicativeOperators.tex}
\input{figures/experiments/operators/equality-operators/equalityOperators.tex}
\input{figures/experiments/operators/conditional-operators/conditionalOperators.tex}

>>>>

\chapter{Experiments}\label{chap:experiments}
Declaration files were generated for existing modules uploaded to the NPM registry. The DefinitelyTyped repository was used as a benchmark. Each one of the generated files was compared against the corresponding declaration file already uploaded to the repository.

\figref{fig:experiments-overall-funnel} shows that a declaration file was generated for 244 modules out of 6029 modules. The following sections elaborate each stage of the process and expose the reasons for generating a declaration file for only 4\% of the modules uploaded to DefinitelyTyped. Samples of the generated declaration files for templates \mintinline{text}{module}, \mintinline{text}{module-class} and \mintinline{text}{module-function} are presented in \secref{sec:experiments-declaration-files-generation}.

\input{figures/experiments/overall-funnel.tex}

Finally, an analysis on the usage of JavaScript operators can be seen in \secref{sec:experiments-js-operators-usage}. The goal is to gather empirical evidence that shows which are the most common types that developers use with every JavaScript operator.

\section{Definitely Typed Benchmark}

\subsection{Code Examples}
Retrieving the code examples from the JavaScript libraries' repositories proved to be a pragmatic way of capturing the types. However, as shown in \figref{fig:experiments-overall-funnel}, working code examples for only 2260 modules could be retrieved. The process of getting a valid code example for a module is divided in 4 blocks:
\begin{itemize}
	\item Extracting repositories url.
	\item Extracting readme files.
	\item Extracting code examples within readme files.
	\item Executing code examples and discarding failing ones.
\end{itemize}

The results obtained for each on of them are described in the following sections.

\subsubsection{Repositories URL}
The url of the repositories could be retrieved for only 4974 modules. More than 1000 modules do not have the repository entry in their corresponding \mintinline{text}{package.json} files. Therefore, the \mintinline{text}{npm view <module> repository.url} command returns an empty value. This is even happening for important modules like \mintinline{text}{ace}.

\subsubsection{Readme Files}
700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \mintinline{text}{readme.md} or \mintinline{text}{README.md}.

\subsubsection{Code Examples Extraction}
The 50\% loss is mainly explained because developers did not wrap their code around a block using the \mintinline{text}{javascript} or \mintinline{text}{js} tags. As explained in \secref{sec:conclusions-future-work}, the process of extracting code examples from a repository can be greatly improved. However, counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.

\subsubsection{Code Examples Execution}
The 2260 extracted code examples were executed by installing the required packages and running the code as a \mintinline{text}{node} application. Working and functional code examples could only be extracted for 1448 modules. 812 modules did not run correctly and were discarded. Some failing samples were analyzed and there were mainly two reasons for the failure:
\begin{enumerate}
	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
\end{enumerate}

Improving the retrieval of code examples for a specific module will definitely have a positive impact in this number.

\subsection{Run-time Information Gathering} \label{sec:experiments-run-time-information-gathering}
Run-time information could only be extracted for 442 modules out of 1448 modules with working code examples. As explained in \secref{sec:run-time-wrapper-objects}, the behavior of the code under analysis was explicitly modified by wrapping the arguments around Wrapper Objects. It definitely had an influence on some modules by causing run-time errors when executing the code examples. Furthermore, Jalangi's instrumentation itself caused some executions to fail, since the modules contained JavaScript features that are not supported by Jalangi. As a result, these modules could not be executed.

An instrumentation without user defined analysis modules was not applied to the 1448 working code examples. Therefore, it was not possible to determine which modules were failing only because of Jalangi's own limitations.

\subsection{Declaration Files Generation} \label{sec:experiments-declaration-files-generation}
Finally, a declaration file was generated for 244 modules. Despite the correct execution of instrumented code examples, some of the 442 modules mentioned in \secref{sec:experiments-run-time-information-gathering} did not contain suitable run-time information for generating a declaration file. 198 of the extracted code examples did not execute the library itself and therefore the collected run-time information was not useful for generating a declaration file.

\subsubsection{Results}
The following section exhibits some samples of the 244 generated declaration files.  It shows some results for each of the implemented templates: \mintinline{text}{module}, \mintinline{text}{module-function} and \mintinline{text}{module-class}.

\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \mintinline{text}{abs}, \mintinline{text}{dirname-regex} and \mintinline{text}{escape-html}. All of them were generated using the \mintinline{text}{module-function} template. There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.

Templates of type \mintinline{text}{module-class} are shown for modules \mintinline{text}{flake-idgen}, \mintinline{text}{route-parser} and \mintinline{text}{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen}, \figref{fig:experiments-results-module-class-route-parser} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation.

Finally, \mintinline{text}{module} template is presented for \mintinline{text}{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct. For example, for \mintinline{text}{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped. However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included. The limitations of using DefinitelyTyped as a baseline for comparison are explained in detail in \chapref{chap:conclusion}.

\input{figures/experiments/results/module-function-results.tex}

\input{figures/experiments/results/module-class-results-flake-idgen.tex}

\input{figures/experiments/results/module-class-results-route-parser.tex}

\input{figures/experiments/results/module-class-results-timer-machine.tex}

\input{figures/experiments/results/module-results.tex}

\input{figures/experiments/results/datadog-metrics.tex}

\subsection{Evaluation}
As shown in \figref{fig:experiments-typescript-templates-distribution-definitely-typed}, 20\% of the declaration files in DefinitelyTyped are written using the \mintinline{text}{module-function}. However, 57\% of the 244 generated declaration files are written with the \mintinline{text}{module-function} template. Additionally, the complexity of evaluating declaration files written with the \mintinline{text}{module-function} is considerably lower than for other templates. The evaluation for templates \mintinline{text}{module-class} and \mintinline{text}{module} was not implemented.

33 out of 116 evaluated modules have no difference with their corresponding declaration file in DefinitelyTyped.

\input{figures/experiments/template-distribution-definitely-typed.tex}

\section{JavaScript Operators Usage} \label{sec:experiments-js-operators-usage}
The following section presents a study on the usage of JavaScript operators. The types of the operands of each operator were recorded with Jalangi. The goal is to better understand how developers are using JavaScript operators and what they expect from them. The study was performed on the 442 modules used for gathering the run-time information, as shown in \figref{fig:experiments-overall-funnel}.

The results might have exposed unexpected development patterns like multiplying a \mintinline{text}{function} and an \mintinline{text}{object}. Conversely, the obtained results support the idea that developers are using operators in a traditional way, despite the possibility in JavaScript of using any type with any operator.

Table \ref{table:experiments-number-of-events-per-operator} summarizes the distribution of events per operator. In total, 1263866 events were collected.

The most relevant results are presented in the following sections.

\input{figures/experiments/operators/operators-distribution.tex}

\subsection{Relational operators}
As shown in \figref{fig:experiments-relational-operators}, operators \mintinline{text}{<, >, <=, >=} are only used with both operands of type \mintinline{text}{number}.

Operator \mintinline{text}{instanceof} is mostly used in the expected way. First and second operand are of type \mintinline{text}{object} and \mintinline{text}{function}, respectively.

Operator \mintinline{text}{in} will check if the second operand has a property that matches the name of the first operand. Most of the events are \mintinline{text}{[number] in [array]}, which means that developers are using the operator to check whether an array has a specific key. The expected usage of this operator is \mintinline{text}{[string] in [object]}, which matches the second most used combination.

\subsection{Additive operators}
Operator \mintinline{text}{+} is mostly used for adding numbers. The combination \mintinline{text}{number-number} represents 77\% of all collected events for this operator. Moreover, 16\% of all events corresponds to the pure string concatenation combination \mintinline{text}{string-string}. Finally, the concatenation of a string a number represents only 1.5\% of all events.

On the other hand, operator \mintinline{text}{-} is only used with type \mintinline{text}{number} for both operands.

Results for additive operators are summarized in \figref{fig:experiments-additive-operators}.

\subsection{Multiplicative operators}
It can be seen in \figref{fig:experiments-multiplicative-operators} that operators \mintinline{text}{*, /, %} are only used with type \mintinline{text}{number}.

\subsection{Equality operators}
\figref{fig:experiments-equality-operators} exhibits the usage of operators \mintinline{text}{==, !=, ===, !==}.

Number and string comparison represent the most used combination for both operators \mintinline{text}{==} and \mintinline{text}{===}.

Operators \mintinline{text}{!=, !==} are mostly used for performing a comparison against \mintinline{text}{null}.

\subsection{Conditional operators}
Jalangi executes the same callback for all operators that trigger a condition check before branching. Therefore, \figref{fig:experiments-conditional-operators} combines operators \mintinline{text}{if-else}, \mintinline{text}{switch-case}, \mintinline{text}{while}, \mintinline{text}{for}, \mintinline{text}{||}, \mintinline{text}{&&} and \mintinline{text}{? :}.

Representing an 82.63\% of all collected events, type \mintinline{text}{boolean} is the most used type with these operators. Types \mintinline{text}{number} and \mintinline{text}{string} come in second and third place, respectively.

\input{figures/experiments/operators/relational-operators/relationalOperators.tex}

\input{figures/experiments/operators/additive-operators/additiveOperators.tex}

\input{figures/experiments/operators/multiplicative-operators/multiplicativeOperators.tex}

\input{figures/experiments/operators/equality-operators/equalityOperators.tex}

\input{figures/experiments/operators/conditional-operators/conditionalOperators.tex}

### 2019-08-18 chapters/6-conclusions.tex
Finish experiments section and conclusion

Brief wrapup of the achievements of this thesis.

Directions for future work (if applicable).

\draft{Una de las conclusiones mas importantes es que la type inference al final no es tan importante. Lo dificil es generar el declaration file como corresponde, detectar las interfaces, etc. El tipo de las variables puede, en principio, inferirse de los ejemplos de los developers. Es una aproximación bastante acertada y se atacan los otros problemas.}

\draft{Decir que todos los bloques pueden mejorarse independientemente: Extraer mejor los readme files, mejorar la instrumentation usando Babel para quedarse siempre en ES5... mejorar la generation de declaration files, etc.}

\section{Future Work}

\subsection{Type Inference}
\todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}

\todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos!!! El concepto de inferir el tipo en realidad pasa a detectar la intencion del developer.}

\subsubsection{Return values}
\todo{El return value se podria incluir en el analysis de Jalangi, detectando las interactions que tiene etc etc. Esto no es tan facil porque no se puede simplemente inspeccionar el return value, sino que hay que saber para qué se usa. Los ejemplos podrían ser útiles}.

\subsubsection{Variable Names}
\todo{Se puede extraer info del nombre de las variables y del nombre de los metodos, etc. La developer's intention.}

\subsection{Uploading Results to Definitely Typed Repository}

>>>>

We have presented \mintinline{text}{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library. The tool downloads code samples written by the developers from the library's repository. It uses these samples to execute the library and gather data flow and type information. The tool finally generates a TypeScript declaration file based on the information gathered at run-time.

We developed an architecture that supports the automatic generation of declaration files for specific JavaScript libraries without additional manual tasks. The architecture contemplates a future incorporation of a Symbolic Execution Engine that refines the initial code base enabling the exploration of new execution paths. However not implemented in this work, its incorporation would result in small incremental modifications to the presented architecture as it is considered to only expand the existing code base.

Building and end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy. Consequently, types were taken over from the values at run-time. Obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation.

We built a mechanism to automatically create declaration files for every module uploaded to DefinitelyTyped and evaluate the results, creating a TypeScript declaration files parser and a comparator.

We exposed the fundamental aspect of capturing the developer's intention when inferring types in JavaScript. Instead of applying constraints and restrictions for operations with certain types, we presented a proposal where common practices are favored. Uncommon usage is not forbidden but greatly disfavored. Used in a first place merely as an initial code base for executing the library and approximating types, code examples turned out to be extremely useful for inferring types. They expose the developer's intention explicitly through code.

Finally, the architecture is composed of different blocks that interact with each other. Each block is independent and has a well defined behavior as well as clear input and output values. As a result, each block can be independently and simultaneously improved.

\section{Future Work} \label{sec:conclusions-future-work}

\subsection{Instrumentation}
\draft{Babel}

\subsection{Declaration files generation}
\draft{Optional parameters}
\draft{Functions as paramters}
\subsubsection{Uploading Results to Definitely Typed Repository}

\subsection{Type Inference}
% \todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}

% \todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos!!! El concepto de inferir el tipo en realidad pasa a detectar la intencion del developer.}

\subsubsection{Return values}
% \todo{El return value se podria incluir en el analysis de Jalangi, detectando las interactions que tiene etc etc. Esto no es tan facil porque no se puede simplemente inspeccionar el return value, sino que hay que saber para qué se usa. Los ejemplos podrían ser útiles}.

\subsubsection{Variable Names}
% \todo{Se puede extraer info del nombre de las variables y del nombre de los metodos, etc. La developer's intention.}

\subsubsection{Variable Content}

\subsubsection{Function Names}

### 2019-08-18 figures/experiments/operators/multiplicative-operators/multiplicativeOperators.tex
Add content to the TypeInference section
Finish experiments section and conclusion

\begin{figure}[h]
	\centering

@@

	\centering
	\hspace{0.1\textwidth}
	\begin{tikzpicture}
		\begin{axis}[
			hide axis,
			scale only axis,
			height=0pt,
			width=0pt,
			colormap/hot,
			colorbar horizontal,
			point meta min=0,
			point meta max=100,
			colorbar style={
				title=\%,
				width=0.7\textwidth,
				xtick={0, 10, 20, 30, 40, ..., 100}
			}]
			\addplot [draw=none] coordinates {(0,0)};
		\end{axis}
	\end{tikzpicture}
	\caption[Relational operators]{\textbf{Type distribution for multiplicative operators} - All multiplicative operators are mainly used with the \textit{number-number} combination}
\end{figure}

>>>>

\begin{figure}[tp]
	\centering

@@

	\centering
	\hspace{0.1\textwidth}
	\begin{tikzpicture}
		\begin{axis}[
			hide axis,
			scale only axis,
			height=0pt,
			width=0pt,
			colormap/hot,
			colorbar horizontal,
			point meta min=0,
			point meta max=100,
			colorbar style={
				title=\%,
				width=0.7\textwidth,
				xtick={0, 10, 20, 30, 40, ..., 100}
			}]
			\addplot [draw=none] coordinates {(0,0)};
		\end{axis}
	\end{tikzpicture}
	\caption[Relational operators]{\textbf{Type distribution for multiplicative operators} - All multiplicative operators are mainly used with the \textit{number-number} combination.}
	\label{fig:experiments-multiplicative-operators}
\end{figure}

### 2019-08-18 figures/experiments/template-distribution-definitely-typed.tex
Finish experiments section and conclusion

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width=0.7\textwidth,
			ybar=0pt,
			ymax=6000,
			enlargelimits=0.15,
			bar width=0.08\textwidth,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={module-class,module-function,module},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=45,anchor=east},
		]
		\addplot coordinates {
			(module-class, 503)
			(module-function, 1201)
			(module, 4379)
		};
		\end{axis}
	\end{tikzpicture}

	\caption[TypeScript templates distribution in DefinitelyTyped]{\textbf{TypeScript templates distribution in DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \mintinline{text}{module} template.
	}

>>>>

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width=0.7\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			bar width=0.08\textwidth,
			legend columns=1,
			legend cell align={left},
			legend style={at={(0.6,0.95)}, anchor=north},
			symbolic x coords={module-class,module-function,module},
			xtick=data,
			ylabel=Percentage,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=45,anchor=east},
		]
		\addplot coordinates {
			(module-class, 8)
			(module-function, 20)
			(module, 72)
		};

		\addplot coordinates {
			(module-class, 12)
			(module-function, 57)
			(module, 31)
		};
		\legend{DefinitelyTyped (Total: 6029), Generated (Total: 244)}
		\end{axis}
	\end{tikzpicture}

	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \mintinline{text}{module} template and only 20\% use the \mintinline{text}{module-function} one. However, 57\% of the 244 generated declaration files use the \mintinline{text}{module-function} template.
	}
